### 2024-03-11 paper/paper.qmd
Draft of codes

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yanyu Wu
  - Ziyi Liu
  - Hechen Zhang
thanks: "Code and data are available at: https://github.com/HechenZ123/Cause-of-Deaths-in-Alberta.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(broom.mixed)
library(tidyverse)
library(janitor)
library(knitr)
library(modelsummary)
library(rstanarm)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
alberta_cod <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
alberta_cod |>
  filter(
    calendar_year == 2022,
    ranking <= 8
  ) |>
  mutate(total_deaths = format(total_deaths, big.mark = ",")) |>
  kable(
    col.names = c("Year", "Cause", "Ranking", "Deaths", "Years"),
    align = c("l", "r", "r", "r", "r"),
    digits = 0, booktabs = TRUE, linesep = ""
  )
```
```{r}
# Look up the five most common causes of death in 2022 of those that have been present every year.
alberta_cod_top_five <-
  alberta_cod |>
  filter(
     calendar_year == 2022
  ) |>
  slice_max(order_by = desc(ranking), n = 5) |>
  pull(cause)
head(alberta_cod_top_five)

alberta_cod <-
  alberta_cod |>
  filter(cause %in% alberta_cod_top_five)
head(alberta_cod)
```

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
alberta_cod |>
  ggplot(aes(x = calendar_year, y = total_deaths, color = cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Annual number of deaths in Alberta") +
  facet_wrap(vars(cause), dir = "v", ncol = 1) +
  theme(legend.position = "none")
```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

cause_of_death_alberta_poisson <-
  readRDS(file = here::here("models/cause_of_death_alberta_poisson.rds"))

cause_of_death_alberta_neg_binomial <-
  readRDS(file = here::here("models/cause_of_death_alberta_neg_binomial.rds"))
```

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
coef_short_names <- 
  c("causeOrganic dementia"
    = "causeOrganic dementia",
    "causeOther ill-defined and unkno..."
    = "causeOther ill-defined...",
    "causeCOVID-19, virus identified"
    = "causeCOVID-19",
    "causeMalignant neoplasms of trac..."
    = "causeMalignant neoplasms of trac..."
    )

modelsummary(
  list(
    "Poisson" = cause_of_death_alberta_poisson,
    "Negative binomial" = cause_of_death_alberta_neg_binomial
  ),
  coef_map = coef_short_names
)

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(cause_of_death_alberta_poisson) +
  theme(legend.position = "bottom")

pp_check(cause_of_death_alberta_neg_binomial) +
  theme(legend.position = "bottom")
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

poisson <- loo(cause_of_death_alberta_poisson, cores = 2)
neg_binomial <- loo(cause_of_death_alberta_neg_binomial, cores = 2)

loo_compare(poisson, neg_binomial)
```

\newpage

# References

### 2024-03-11 scripts/02-data_cleaning.R
Update 02-data_cleaning and analysis_data.csv

@@ -10,7 +10,7 @@ library(knitr)
 
 #### Clean data ####
 # Look at the distribution of these deaths, by year and cause.
-alberta_cod <-
+clean_data <-
   read_csv(
     "data/raw_data/raw-data-deaths-leading-causes.csv",
     skip = 2,
@@ -25,32 +25,8 @@ alberta_cod <-
   add_count(cause) |>
   mutate(cause = str_trunc(cause, 30))
 
-# look at the top-eight causes in 2022
-alberta_cod |>
-  filter(
-    calendar_year == 2022,
-    ranking <= 8
-  ) |>
-  mutate(total_deaths = format(total_deaths, big.mark = ",")) |>
-  kable(
-    col.names = c("Year", "Cause", "Ranking", "Deaths", "Years"),
-    align = c("l", "r", "r", "r", "r"),
-    digits = 0, booktabs = TRUE, linesep = ""
-  )
-
-# Look up the five most common causes of death in 2022 of those that have been present every year.
-alberta_cod_top_five <-
-  alberta_cod |>
-  filter(
-    calendar_year == 2022
-  ) |>
-  slice_max(order_by = desc(ranking), n = 5) |>
-  pull(cause)
-alberta_cod_top_five
-
-alberta_cod <-
-  alberta_cod |>
-  filter(cause %in% alberta_cod_top_five)
+# Drop rows with missing data
+clean_data <- na.omit(clean_data)
 
 #### Save data ####
-write_csv(alberta_cod, "data/analysis_data/analysis_data.csv")
+write_csv(clean_data, "data/analysis_data/analysis_data.csv")

### 2024-03-11 scripts/03-test_data.R
Finished 03-test_data

@@ -1,15 +1,58 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Tests Simulated Data and Analysis Data
+# Date: 11 March 2024
+# Contact: hechen.zhang@mail.utoronto.ca
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+#### Test Simulated Data ####
+# Load the simulated dataset
+simulated_data <- read.csv("data/raw_data/simulated_data.csv")
+
+# Check if 'cause' is character
+is.character(simulated_data$cause)
+
+# Check if 'year' is numeric
+is.numeric(simulated_data$year)
+
+# Check if 'deaths' is numeric
+is.numeric(simulated_data$deaths)
+
+# Check if any numerical data is mistakenly entered as character, or any character data is mistakenly entered as number.
+!is.numeric(simulated_data$cause)
+!is.character(simulated_data$year)
+!is.character(simulated_data$deaths)
+
+# Check if categorical data variables contain expected unique values only
+all(simulated_data$cause %in% c('Heart', 'Stroke', 'Diabetes'))
+
+
+#### Test Real Data ####
+# Load analysis_data.csv
+analysis_data <- read.csv("data/analysis_data/analysis_data.csv")
+
+library(readr)
+
+# Test data types
+is.integer(analysis_data$calendar_year)
+is.character(analysis_data$cause)
+is.integer(analysis_data$ranking)
+is.integer(analysis_data$total_deaths)
+is.integer(analysis_data$n)
+
+# Verify calendar_year range
+analysis_data$calendar_year
+all(analysis_data$calendar_year >= 2001 & analysis_data$calendar_year <= 2022)
+
+# Ensure ranking is positive
+all(analysis_data$ranking > 0 & analysis_data$ranking <= 100)
+
+# Ensure total_deaths is non-negative
+all(analysis_data$total_deaths >= 0)
+
+# Check 'n' for specific range 
+all(analysis_data$n >= 1 & analysis_data$n <= 1000)

### 2024-03-12 paper/datasheet_template.qmd
Fill out some of the questions

---
title: "Datasheet for 'A dataset'"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: https://github.com/HechenZ123/Cause-of-Deaths-in-Alberta.git"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - To close gaps in complete mortality data and facilitate analysis of health trends, this dataset was created to examine the leading causes of death in Alberta each year. It provides detailed cause of death information to support policy development. 
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Service Alberta created the dataset on behalf of Alberta Government.

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The dataset of leading causes of death in Alberta was released by the Government of Alberta
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance in the dataset represents a specific leading cause of death for a given calendar year. This setup implies that the dataset is a collection of records, each documenting the frequency of deaths attributed to various causes in different years. The dataset focuses on 8 type of instances which are causes of death: Organic dementia, All other forms of chronic ischemic heart disease, Other ill-defined and unknown causes of mortality, COVID-19, virus identified, Malignant neoplasms of trachea, bronchus and lung, Acute myocardial infarction, Accidental poisoning by and exposure to drugs and other biological substances, Other chronic obstructive pulmonary disease. It also captures multiple dimensions of information such as temporal and quantitative. All data points revolve around the central theme of mortality causes.

2. *How many instances are there in total (of each type, if appropriate)?*
	- There are 8 instances in total.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- This dataset lists the leading causes of death in Alberta, covering the 30 most common causes each year. Given this structure, it does not appear to be a sample from a larger dataset, but rather a comprehensive list within its definition (i.e., the top 30 causes of death). This suggests that the dataset is designed to capture the most important aspects of mortality within the province without the need for a random sample from a larger dataset. However, because it focuses on the top causes of death, it may not be able to cover all possible causes of death, thus excluding less common causes. This limitation means that while the dataset represents only the most influential issues contributing to mortality within Alberta, it may not be fully representative of all causes of death.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- The labels associated with each instance are a total number of deaths for each instance and the ranking of each instance (the most common causes of death in Alberta).

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-03-12 paper/paper.qmd
Updated the latest version

@@
 thanks: "Code and data are available at: https://github.com/HechenZ123/Cause-of-Deaths-in-Alberta.git"
-date: today
-date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+date: 'today'
+date-format: 'long'
+abstract: "When the continuity assumption of linear regression does not apply, we can work with counting data using Poisson regression, which deals with discontinuous results but still incorporates predictor variables into the model in a linear manner.
+We demonstrate the validity of negative binomial regression through an extensive analysis of mortality in Alberta, Canada, where the data is too scattered for the classical method (Poisson regression).
+Based on our findings, negative binomial regression improves our ability to predict outcomes by fitting the data more precisely, especially in cases where the variance is substantially higher than the mean.
+This study highlights how important it is to use the right statistical models to get more accurate findings, which will ultimately enhance our understanding of patterns and trends in many academic fields."
 format: pdf
 toc: true
-number-sections: true
+editor: visual
+render:
+  latex_engine: xelatex
 bibliography: references.bib
 ---
+\newpage
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
+# 1.0 Introduction
+
+The mortality rate, often referred to as the death rate, represents an approximation of the fraction of a population that dies within a given time frame [@porta2014dictionary].
+Mortality rates can serve as a crucial indicator of a population’s health status, and it also reveals the impact of diseases and other health-related issues over a period of time.
+This paper explores the leading causes of death in Alberta for crafting effective public health strategies and policies and understanding the most significant health threats affecting a population for researchers[@datasetCOD].
+By identifying the main causes of mortality, health authorities can prioritise research funding towards diseases and conditions that have the highest impact on community health and lifespan [@vargas2019assessment].
+
+As discussed in the data section, we used data from Service Alberta [@datasetCOD] on the leading causes of deaths, in which the five most significant causes in 2022 were analysed.
+These five causes are Organic dementia, Other causes not clearly defined, COVID-19, and Cancers of the trachea, bronchus, and lungs.
+It was noted that, among the examples mentioned, the negative binomial regression is more accurate compared to the Poisson model, while Poisson regression is prone to errors.
+
+## 1.1 Importing Important Packages.
+In this analysis, we employ a range of R [@citeR] packages tailored for data cleaning, transformation, analysis, and reporting.
+`Tidyverse` by @tidyverse is used for data wrangling, `janitor` package by @janitor is used for data cleaning operations, and `knitr` by @knitr for data presentation in data tables.The following code section aims at importing the important packages that are essential for examining the missing values in the data set.We run the model in R @citeR using the `rstanarm` package of @rstanarm.
+We use the default priors from `rstanarm`.
+For comprehensive mixed effects model analysis, we leverage the `broom.mixed` package [@broommixed], which extends the `broom` package functionalities to mixed models, facilitating the extraction, tidying, and representation of model outputs.
+Furthermore, the `modelsummary` package [@modelsummary] provides tools for creating customizable summary tables of model results, enhancing the interpretability and dissemination of statistical findings.
+By calculating the LOO-CV scores for different models with `loo` [@loo], we could compare them based on their out-of-sample predictive accuracy.
+Lower values of LOOIC indicate better model performance.
+The following code sections aim to import these crucial packages, essential for conducting a thorough analysis and addressing the research questions at hand, while ensuring data integrity and transparent reporting of results.
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+# Set up libraries
 library(broom.mixed)
@@
 library(rstanarm)
+library(loo)
 ```
 
-# Introduction
+## 1.2. Data Overview.
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Our data is of leading causes of death (@fig-dataset-overview), from @datasetCOD.
 
-The remainder of this paper is structured as follows. @sec-data....
-
-# Data {#sec-data}
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Examining the top ten causes in 2021 reveals several notable findings (@fig-dataset-overview). For example, ...
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-dataset-overview
+#| fig-cap: Top-teight causes of death in Alberta in 2022
 #| echo: false
+#| warning: false
 alberta_cod <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
@@
 ```
+
+# 2.0 Data {#sec-data}
+
+For simplicity we restrict ourselves to the five most common causes of death in 2022 of those that have been present every year.
 ```{r}
-# Look up the five most common causes of death in 2022 of those that have been present every year.
+#| label: fig-top-five
+#| fig-cap: Top Five Leading causes of death in 2022
+#| echo: false
+#| warning: false
+# Look up the five most common causes of death in 2022
 alberta_cod_top_five <-
@@
   filter(cause %in% alberta_cod_top_five)
-head(alberta_cod)
 ```
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 #| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| fig-cap:  Annual number of deaths for the top-five causes in 2022, since 2001, for Alberta, Canada
 #| echo: false
@@
 
-# Model
+```{r}
+#| label: fig-summary_table
+#| echo: false
+#| fig-cap:   Summary statistics of the number of yearly deaths, by cause, in Alberta, Canada
+#| warning: false
+#| message: false
+summary(alberta_cod$total_deaths)
+```
+
+# 3.0 Model
 
@@
 
-## Model set-up
+## 3.1 Model set-up
 
@@
 #| eval: true
+#| label: tbl-models
 #| warning: false
-#| message: false
 
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: "Modeling the most prevalent cause of deaths in Alberta, 2001-2022"
 #| warning: false
-coef_short_names <- 
-  c("causeOrganic dementia"
-    = "causeOrganic dementia",
-    "causeOther ill-defined and unkno..."
-    = "causeOther ill-defined...",
-    "causeCOVID-19, virus identified"
-    = "causeCOVID-19",
-    "causeMalignant neoplasms of trac..."
-    = "causeMalignant neoplasms of trac..."
-    )
-
 modelsummary(
@@
     "Negative binomial" = cause_of_death_alberta_neg_binomial
-  ),
-  coef_map = coef_short_names
+  )
 )
@@
 
-## First discussion point {#sec-first-point}
+## Addressing Public Health Challenges in Alberta: Strategies for Health Policy and Social Regulation
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Based on the top five causes of death in Alberta, it is imperative that we consider initiatives to improve health policy and social regulation to address health challenges, prioritizing public health issues.
+Looking at the data, given the significant impact of Covid-19 on mortality rates, it is crucial to continue efforts to implement public health interventions.
+This includes increasing mass vaccination activities, continuing to promote mask-wearing and social distancing measures, and enhancing testing and contact tracing capabilities.
+Furthermore, it is essential to ensure that the healthcare system has sufficient capacity to handle an increase in cases.
+It is worth noting that in the coming years, due to the passage of several years since the virus initially emerged, Covid-19 may not continue to be such a significant cause of death, as the virus gradually becomes less virulent or severe [@talic2021effectiveness].
+Implementing policies for cancer prevention and control can help reduce mortality from malignant neoplasms of the trachea, bronchus, and lung.
+This may include implementing tobacco control measures, such as increasing tobacco product taxes, comprehensive smoking cessation programs, and restricting tobacco advertising and promotion.
+Additionally, promoting healthy lifestyles, early cancer screening programs, and providing high-quality cancer treatment services are also crucial [@eastman2023nci].
 
@@
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-cap: "Comparing posterior prediction checks for Poisson and negative binomial models"
+#| fig-subcap: ["Poisson model", "Negative binomial model"]
 
@@
 
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
 ```{r}
@@
 #| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| message: true
+#| warning: true
+#| label: fig-LOOIC
+#| fig-cap: "Comparing the LOO-CV scores"
 

### 2024-03-15 paper/paper.qmd
Updated

@@
 
-## 1.2. Data Overview.
+# 2.0 Data {#sec-data}
 
@@
 
-Examining the top ten causes in 2021 reveals several notable findings (@fig-dataset-overview). For example, ...
+Examining the top ten causes in 2022 reveals several notable findings (@fig-dataset-overview). For example, ...
 
@@
 ```
-
-# 2.0 Data {#sec-data}
 
@@
 #| message: false
-summary(alberta_cod$total_deaths)
+# Calculate summary statistics for total_deaths
+summary_stats <- data.frame(
+  Min = min(alberta_cod$total_deaths, na.rm = TRUE),
+  Mean = mean(alberta_cod$total_deaths, na.rm = TRUE),
+  Max = max(alberta_cod$total_deaths, na.rm = TRUE),
+  SD = sd(alberta_cod$total_deaths, na.rm = TRUE),
+  Var = var(alberta_cod$total_deaths, na.rm = TRUE),
+  N = sum(!is.na(alberta_cod$total_deaths))
+)
+
+# Print the summary statistics
+print(summary_stats)
+
 ```
@@
 
-poisson <- loo(cause_of_death_alberta_poisson, cores = 2)
-neg_binomial <- loo(cause_of_death_alberta_neg_binomial, cores = 2)
+poisson <- loo(cause_of_death_alberta_poisson, cores = 1)
+neg_binomial <- loo(cause_of_death_alberta_neg_binomial, cores = 1)
 

### 2024-03-16 paper/paper.qmd
Update Data section

--- 2024-03-12 paper/paper.qmd
@@
 date-format: 'long'
-abstract: "When the continuity assumption of linear regression does not apply, we can work with counting data using Poisson regression, which deals with discontinuous results but still incorporates predictor variables into the model in a linear manner.
-We demonstrate the validity of negative binomial regression through an extensive analysis of mortality in Alberta, Canada, where the data is too scattered for the classical method (Poisson regression).
-Based on our findings, negative binomial regression improves our ability to predict outcomes by fitting the data more precisely, especially in cases where the variance is substantially higher than the mean.
-This study highlights how important it is to use the right statistical models to get more accurate findings, which will ultimately enhance our understanding of patterns and trends in many academic fields."
+abstract: "We examined the distribution of the top 30 causes of death for each year between 2001 and 2022 in Alberta, Canada.
+The conifer model and negative binomial regression were used to analyse the long-term leading causes of death and the sudden emergence of specific causes of death.
+According to our findings, negative binomial regression improves our ability to predict outcomes when the data is too spread out by fitting the data more accurately.
+The results could not only help policymakers design more effective preventive measures to reduce mortality from these conditions, but also help researchers and policymakers make more precise decisions."
 format: pdf
@@
 
-Examining the top ten causes in 2021 reveals several notable findings (@fig-dataset-overview). For example, ...
+## 2.1 Data Sources
+The Alberta Government created the dataset.The official website of the Government of Alberta provides a single platform for services and information pertaining to Alberta, Canada.
+It provides information on government news, jobs, social services, health, and education, making it an essential tool for both locals and tourists to be updated about provincial issues.
+
+## 2.2 Variables
+Order by total number of deaths and a ranking of the top 30 causes of death in Alberta each year.Our table lists the top eight causes of death in Alberta in 2022. Examine each variable in detail:
+*Year*: This denotes the data gathering year, which for all entries is 2022.
+*Cause*: This represents the medical condition or event that led to death. The causes listed are shown in @fig-dataset-overview:
+`Organic dementia`
+`All other forms of chronic...`, 
+`Other ill-defined and unknown...`, 
+`COVID-19, virus identified`, 
+`Malignant neoplasms of the trachea, bronchus, and lung`, 
+`Acute myocardial infarction`, 
+`Accidental poisoning by and...`, 
+`Other chronic obstructive pulmonary diseases`, 
+*Ranking*: This is a ranking by the number of deaths caused by each disease, with 1 being the highest.
+*Deaths*: The number of deaths attributed to each cause.
+*Year*: Indicates the number of years in which data was collected for that reason.
 
@@
 
-# 2.0 Data {#sec-data}
+## 2.3 Data Description
+It accurately reflects health issues specific to the region, but this information is somewhat limited by location.
+Health policies vary from country to country, and in some regions and countries, health coverage may be less developed than in others, and low-income groups may not be able to afford basic medical care, which increases their risk of chronic diseases and other health problems.
+The main types of employment in a city also have an impact on health; In an industrial city, some jobs may require exposure to hazardous substances or strenuous physical labor, which increases the likelihood of developing occupational diseases.
+Risks associated with the workplace may lead to an increase in diseases such as cardiovascular and respiratory diseases.
+In addition, communities with higher levels of education and economic prosperity generally have stronger social support systems, which are critical for disease prevention.
+
+The demographics of a society also have a significant impact on cause-of-death data.
+Age distribution is important because certain age groups are more susceptible to certain diseases.
+For example, young children may be more susceptible to infectious diseases, while older people may be more susceptible to degenerative and chronic diseases.
+Given that certain medical conditions are more common in one sex than the other - men may have higher rates of heart disease, while women may have higher rates of certain cancers - the sex ratio may also have an impact on causes of death.
+In addition, differences in the ethnic makeup of the population may be related to lifestyle choices, genetic makeup, and disease risk.
+For example, genetic factors may contribute to the higher incidence of certain chronic diseases in particular ethnic groups.
+Elucidating how resources are allocated to public health problems therefore requires a thorough understanding and study of population characteristics.
+
+We can also see that in the data table @fig-dataset-overview for 2022, COVID-19 ranks fourth, with 1,547 deaths.
+Alberta's cause-of-death data proves that the coronavirus pandemic is one of the most serious public health crises of the early 21st century, with global implications.
+This figure not only speaks to the outbreak's death rate, but also suggests that the region's health care system may be under strain.
+The COVID-19 pandemic has disrupted the distribution of deaths from previously common causes, which may include some chronic diseases that have long held the top spot, such as cancer and heart disease, even though it was not the leading cause of death in the years shown in the table.
+
+As large datasets are difficult to observe, this report will analyze only specific aspects.
+The original dataset contains eight different causes of death.
+Only the first five causes will be analyzed in this paper.
 
@@
 ```{r}
-#| label: fig-top-five
-#| fig-cap: Top Five Leading causes of death in 2022
+#| label: tbl-top-five
+#| tbl-cap: Top Five Leading causes of death in 2022
 #| echo: false
@@
   pull(cause)
-head(alberta_cod_top_five)
 
@@
   filter(cause %in% alberta_cod_top_five)
-```
 
-```{r}
-#| label: fig-planes
-#| fig-cap:  Annual number of deaths for the top-five causes in 2022, since 2001, for Alberta, Canada
-#| echo: false
-#| warning: false
-#| message: false
-alberta_cod |>
-  ggplot(aes(x = calendar_year, y = total_deaths, color = cause)) +
-  geom_line() +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  labs(x = "Year", y = "Annual number of deaths in Alberta") +
-  facet_wrap(vars(cause), dir = "v", ncol = 1) +
-  theme(legend.position = "none")
-```
-
-```{r}
-#| label: fig-summary_table
-#| echo: false
-#| fig-cap:   Summary statistics of the number of yearly deaths, by cause, in Alberta, Canada
-#| warning: false
-#| message: false
-summary(alberta_cod$total_deaths)
+head(alberta_cod_top_five)
 ```
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in Appendix
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the number of deaths recorded in Alberta. Then $\beta_i$ is ... and $\gamma_i$ is .. $x_i$ is the causes of death.
 
+Poisson model:
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i|\lambda_i &\sim \mbox{Poisson}(\lambda_i) \\
+log(\lambda_i) & = \beta_0 + \beta_1 \cdot x_i\\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\end{align}
+
+Negative binomial model:
+\begin{align} 
+y_i|\lambda_i, \theta &\sim \mbox{NegativeBinomial}(\mu_i, \theta) \\
+log(\mu_i) & = \beta_0 + \beta_1 \cdot x_i\\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
 \end{align}
@@
 # Results
+```{r}
+#| label: fig-planes
+#| fig-cap:  Annual number of deaths for the top-five causes in 2022, since 2001, for Alberta, Canada
+#| echo: false
+#| warning: false
+#| message: false
+alberta_cod |>
+  ggplot(aes(x = calendar_year, y = total_deaths, color = cause)) +
+  geom_line() +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  labs(x = "Year", y = "Annual number of deaths in Alberta") +
+  facet_wrap(vars(cause), dir = "v", ncol = 1) +
+  theme(legend.position = "none")
+```
+
+```{r}
+#| label: fig-summary_table
+#| echo: false
+#| tbl-cap: Summary by cause of the number of yearly deaths in Alberta, Canada
+#| warning: false
+#| message: false
+# Calculate summary statistics for total_deaths
+summary_stats <- data.frame(
+  Min = min(alberta_cod$total_deaths, na.rm = TRUE),
+  Mean = mean(alberta_cod$total_deaths, na.rm = TRUE),
+  Max = max(alberta_cod$total_deaths, na.rm = TRUE),
+  SD = sd(alberta_cod$total_deaths, na.rm = TRUE),
+  Var = var(alberta_cod$total_deaths, na.rm = TRUE),
+  N = sum(!is.na(alberta_cod$total_deaths))
+)
+
+# Use kable from knitr to create a table
+kable(summary_stats)
+```
 

### 2024-03-16 paper/paper.qmd
Model updated

@@
   latex_engine: xelatex
+number-sections: true
 bibliography: references.bib
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+## 3.1 Model setup:
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in Appendix
+In analyzing the association between the total number of deaths and significant causes of death, we used two different regression models to predict, namely the Poisson distribution and the binomial distribution.
+Both the Poisson and binomial distribution models are regression models used to analyze count data, primarily for finding the relationship between independent variables (predictors) and dependent variables (count outcomes).
+In regression models, the independent variables are independent, while the dependent variables depend on the independent variables.
+In our model, the independent variables are the different causes of death mentioned in the table, such as COVID-19, malignant neoplasms, organic dementia, etc.
+These variables are used to predict or explain the number of deaths.
+The dependent variable is the number of deaths in Alberta from 2001 to 2022.
+This variable is the response variable in the model, and its count is the result to be predicted or explained.
 
-## 3.1 Model set-up
+Using the Poisson distribution has different advantages and disadvantages in this study.
+The Poisson model is simple in form, with fewer parameters, making it easy to understand and explain, and because of its simplicity, this model has a faster computation speed and is easy to implement.
+The Poisson distribution has a solid theoretical foundation for modeling the number of times an event occurs.
+It assumes that the average frequency of events occurring over a certain time or space is constant, and the occurrence of individual events is independent.
+At the same time, the Poisson distribution is a regression model specifically used for analyzing count data (where the response variable is non-negative).
+In our study, our response variable is the number of deaths, which is definitely a positive number, making the Poisson distribution suitable for our research.
+However, it also has some shortcomings, such as the problem of overdispersion; if the actual data show that the variance is greater than the mean, the Poisson model may not be applicable, and the model's predictive effect will worsen.
 
-Define $y_i$ as the number of deaths recorded in Alberta. Then $\beta_i$ is ... and $\gamma_i$ is .. $x_i$ is the causes of death.
+Regarding the negative binomial distribution, its advantage is that it includes an additional parameter to model overdispersion, making it more suitable for data where the variance is greater than the mean.
+Compared to the Poisson model, the negative binomial model offers more flexibility to fit various data, especially those that exhibit significant overdispersion or clustering.
+The negative binomial model includes an extra parameter to account for overdispersion, allowing the variance to exceed the mean.
+In our study, this model also attempts to link death counts with various causes but is more flexible in handling data variability.
+However, compared to the Poisson distribution, the negative binomial model is more complex, involving more parameters, which may lead to difficulties in model explanation and communication.
+The negative binomial model typically requires more computational resources, so it may not be as efficient as the Poisson model on large datasets.
+
+In our data situation, the negative binomial model seems more appropriate than the Poisson model, possibly due to overdispersion in the data.
+Therefore, in analyzing causes of death (such as COVID-19, malignant neoplasms, organic dementia, etc.), choosing the negative binomial model may provide more accurate estimates.
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+From this formula, we can consider $y_i$ to be the total number of deaths, with the i-th observation representing the cause of death.
+$\beta_0$ represents the intercept, which is the expected log count of the response variable when all predictor variables are held at zero.
+The intercepts of the two models are similar.
+Each cause of death has a corresponding coefficient $\beta_i$, indicating the change in the expected log count of deaths for each unit increase in the predictor variable (i.e., the presence of the cause of death).
+A negative coefficient suggests that the presence of that cause is associated with a decrease in the death count relative to the baseline level.
+The intercept $\beta_0$ and the slope $\beta_i$ both follow a prior distribution that is normally distributed with a mean of 0 and a standard deviation of 2.5.
 
@@
 #| eval: true
-#| label: tbl-models
 #| warning: false
@@
 
-### Model justification
+## 3.2 Model justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Based on the models discussed above, it is challenging to determine whether each cause of death has a positive or negative correlation with the total number of deaths, as the leading cause of death changes each year.
+The specific impact of these changes must be determined by the coefficient of each leading cause of death to decide whether the correlation is positive or negative.
+According to Table 1, from 2001 to 2022, organic dementia shows an upward trend, indicating that organic dementia is positively correlated with the total number of deaths.
+In the upcoming model results section, we will elaborate in detail on the association between each cause of death and the total number of deaths.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Based on our discussion in the model set-up section, the negative binomial distribution seems more suitable for our research objectives than the Poisson distribution.
+The Poisson model assumes that the mean and variance are equal, but in the real world, they are rarely equal.
+The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.
 

### 2024-03-16 paper/paper.qmd
updated model section

@@
 
-# 1.0 Introduction
+# Introduction
 
@@
 
-## 1.1 Importing Important Packages.
+## Importing Important Packages.
 In this analysis, we employ a range of R [@citeR] packages tailored for data cleaning, transformation, analysis, and reporting.
@@
 
-# 2.0 Data {#sec-data}
+# Data {#sec-data}
 
@@
 
-## 2.1 Data Sources
+## Data Sources
 The Alberta Government created the dataset.The official website of the Government of Alberta provides a single platform for services and information pertaining to Alberta, Canada.
@@
 
-## 2.2 Variables
+## Variables
 Order by total number of deaths and a ranking of the top 30 causes of death in Alberta each year.Our table lists the top eight causes of death in Alberta in 2022. Examine each variable in detail:
@@
 
-## 2.3 Data Description
+## Data Description
 It accurately reflects health issues specific to the region, but this information is somewhat limited by location.
@@
 
-# 3.0 Model
+# Model
 
-## 3.1 Model setup:
+## Model setup:
 
@@
 
-## 3.2 Model justification
+## Model justification
 
@@
 The specific impact of these changes must be determined by the coefficient of each leading cause of death to decide whether the correlation is positive or negative.
-According to Table 1, from 2001 to 2022, organic dementia shows an upward trend, indicating that organic dementia is positively correlated with the total number of deaths.
+According to @fig-planes, from 2001 to 2022, organic dementia shows an upward trend, indicating that organic dementia is positively correlated with the total number of deaths.
 In the upcoming model results section, we will elaborate in detail on the association between each cause of death and the total number of deaths.
@@
 The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.
+
+```{r}
+#| label: tbl-mean-var
+#| tbl-cap: Comparison of Mean and Variance of Total Death in from 2001 to 2022
+#| echo: false
+#| warning: false
+# Calculate the mean and variance
+summary_stats <- alberta_cod %>%
+  summarise(
+    Mean = mean(total_deaths, na.rm = TRUE),  # Calculate mean, remove NA values if any
+    Variance = var(total_deaths, na.rm = TRUE) # Calculate variance, remove NA values if any
+  )
+
+# Create a table
+kable(summary_stats, caption = "Comparison of Mean and Variance of Total Deaths")
+
+```
 

### 2024-03-16 paper/paper.qmd
Model updated

@@
 The Poisson model assumes that the mean and variance are equal, but in the real world, they are rarely equal.
+As @tbl-mean-var shown below indicates, the mean and variance are not equal.
 The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.
@@
 ```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Comparing posterior prediction checks for Poisson and negative binomial models"
+#| fig-subcap: ["Poisson model", "Negative binomial model"]
+
+pp_check(cause_of_death_alberta_poisson) +
+  theme(legend.position = "bottom")
+
+pp_check(cause_of_death_alberta_neg_binomial) +
+  theme(legend.position = "bottom")
+```
+
+We created a figure @fig-ppcheckandposteriorvsprior that includes both the Poisson and negative binomial models.
+For the Poisson model, the actual observed values y are represented in black.
+The predicted values $y_{Rep}$ are shown in light blue, which are replicated samples generated based on the posterior distribution of model parameters.
+
+It can be seen that there is a significant overlap between many of the predictive distributions and the actual observed values, but at higher values, the range of the predictive distribution becomes very wide, indicating that the model may not fit well for high-count values.
+
+For the negative binomial models, the actual observed values y are also represented in black, and the predicted values yrep are likewise shown in light blue, generated based on the posterior distribution of the model parameters.
+Compared to the Poisson model, the predictive distributions generated by the negative binomial model appear to be more concentrated, meaning they are more closely clustered around the actual observed values at higher counts, indicating that the negative binomial model may be more suitable for these data.
+From the figure, it can be observed that the predictive distributions of the negative binomial model match the actual observed values better than those of the Poisson model.
+In the Poisson model's figure, the predicted values align well with the actual values near the peak in the middle but are wider in the tail distribution, which may mean that the Poisson model struggles with extreme values.
+In contrast, in the figure for the negative binomial model, the predicted values seem to follow the actual observed values more closely across the entire range, typically indicating it can better handle the overdispersion of the data.
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: true
+#| warning: true
+#| label: fig-LOOIC
+#| fig-cap: "Comparing the LOO-CV scores"
+
+poisson <- loo(cause_of_death_alberta_poisson, cores = 1)
+neg_binomial <- loo(cause_of_death_alberta_neg_binomial, cores = 1)
+
+loo_compare(poisson, neg_binomial)
+```
+
+To demonstrate that the negative binomial model is a better fit, we compared the Expected Log Predictive Density (ELPD) difference and the Standard Error (SE) difference between these two models.
+The ELPD difference column compares the difference in expected log pointwise predictive density between the two models—the negative binomial model and the Poisson model.
+ELPD is an indicator of model predictive performance, with higher ELPD values typically indicating better model predictive performance.
+For the negative binomial model (listed as cause_of_death_alberta_neg_binomial), the ELPD difference is 0.0.
+For the Poisson model (listed as cause_of_death_alberta_poisson), the ELPD difference is -6160.6, meaning that, relative to the negative binomial model, the Poisson model's predictive performance is worse.
+
+The SE difference represents the uncertainty in the estimation of the ELPD difference.
+A smaller SE difference indicates that the estimation of the ELPD difference is more precise; a larger SE difference indicates that the estimation of the ELPD difference is less stable and has more uncertainty.
+In other words, it tells us the reliability of the difference in predictive performance between the models.
+The SE difference is 1412.1, which is a relatively large value, indicating considerable uncertainty in the estimation of the predictive performance difference between the negative binomial model and the Poisson model.
+Nonetheless, the significant difference in ELPD (-6160.6) compared to its SE suggests that this difference is statistically significant.
+In other words, even considering the uncertainty in the ELPD difference, the negative binomial model's predictive performance is significantly better than that of the Poisson model, supporting our previous statements.
 
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Comparing posterior prediction checks for Poisson and negative binomial models"
-#| fig-subcap: ["Poisson model", "Negative binomial model"]
-
-pp_check(cause_of_death_alberta_poisson) +
-  theme(legend.position = "bottom")
-
-pp_check(cause_of_death_alberta_neg_binomial) +
-  theme(legend.position = "bottom")
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: true
-#| warning: true
-#| label: fig-LOOIC
-#| fig-cap: "Comparing the LOO-CV scores"
-
-poisson <- loo(cause_of_death_alberta_poisson, cores = 1)
-neg_binomial <- loo(cause_of_death_alberta_neg_binomial, cores = 1)
-
-loo_compare(poisson, neg_binomial)
-```
-
 \newpage

### 2024-03-17 paper/paper.qmd
Updated data, model section

@@
 ## Data Sources
-The Alberta Government created the dataset.The official website of the Government of Alberta provides a single platform for services and information pertaining to Alberta, Canada.
-It provides information on government news, jobs, social services, health, and education, making it an essential tool for both locals and tourists to be updated about provincial issues.
+Our research team conducted an in-depth analysis of mortality trends in Alberta, choosing a dataset published on the Alberta Open Government Portal as the basis for their study.
+This dataset was made available to the public beginning in 2015 and was last updated in 2023, and is designed to promote public engagement by providing transparent access to key health statistics.This dataset provides a comprehensive view of deaths across the province, covering the top 30 leading causes of death in each annual count.
+This dataset was chosen because it provides key statistics on the number of deaths and mortality rates.
+These statistics cover data from chronic diseases to unintended injuries and even public health emergencies, allowing us to accurately track trends in key causes of death.
+We chose to rely on this dataset primarily because it is sourced from the official government of Alberta, which provides an authoritative data base for our study.
+The endorsement of a government department ensures the accuracy of the data, which is a key prerequisite for conducting scientific research.
+In addition, the high degree of fit of the dataset with our research objectives was an important reason for its selection.
+Our study aimed to gain insight into the key factors that influence mortality in Alberta and to explore possible public health interventions.
+By analysing this data, we are able to identify priority areas for health interventions and provide a scientific basis for policy makers.
 
-## Variables
+## Features
 Order by total number of deaths and a ranking of the top 30 causes of death in Alberta each year.Our table lists the top eight causes of death in Alberta in 2022. Examine each variable in detail:
@@
 *Deaths*: The number of deaths attributed to each cause.
-*Year*: Indicates the number of years in which data was collected for that reason.
+*Years*: Indicates the number of years in which data was collected for that reason.
 
@@
 
-## Data Description
-It accurately reflects health issues specific to the region, but this information is somewhat limited by location.
-Health policies vary from country to country, and in some regions and countries, health coverage may be less developed than in others, and low-income groups may not be able to afford basic medical care, which increases their risk of chronic diseases and other health problems.
-The main types of employment in a city also have an impact on health; In an industrial city, some jobs may require exposure to hazardous substances or strenuous physical labor, which increases the likelihood of developing occupational diseases.
-Risks associated with the workplace may lead to an increase in diseases such as cardiovascular and respiratory diseases.
-In addition, communities with higher levels of education and economic prosperity generally have stronger social support systems, which are critical for disease prevention.
+## Data Method
+In the preliminary phase, we worked on organizing and harmonizing the naming of variables in the dataset to enhance its readability and efficiency of analysis.
+In addition, we summarized the data in order to quantify the incidence of various causes of death, thus making the dataset more streamlined for more in-depth analysis.
+To avoid missing years in the dataset from affecting the results of the analysis, we decided to delete all records containing these missing values.We then prioritized the top 8 causes of death by 2022 based on their number of deaths.
+Particular attention was paid to the top 5 leading causes of death, which is at a critical time when the world is facing a pandemic virus, and the data allow us to understand the impact of the new crown on mortality.Our mortality data are consistent with the characteristic of occurring independently with equal means and variances.
+These mortality trends are fitted by Poisson and negative binomial statistical models because of the suitability of both models for working with count data.
+A method for assessing the effect of predictor variables on mortality changes is provided, with exponential coefficients that visually represent the change in mortality risk that may result from each unit change in a predictor variable.
 
-The demographics of a society also have a significant impact on cause-of-death data.
-Age distribution is important because certain age groups are more susceptible to certain diseases.
-For example, young children may be more susceptible to infectious diseases, while older people may be more susceptible to degenerative and chronic diseases.
-Given that certain medical conditions are more common in one sex than the other - men may have higher rates of heart disease, while women may have higher rates of certain cancers - the sex ratio may also have an impact on causes of death.
-In addition, differences in the ethnic makeup of the population may be related to lifestyle choices, genetic makeup, and disease risk.
-For example, genetic factors may contribute to the higher incidence of certain chronic diseases in particular ethnic groups.
-Elucidating how resources are allocated to public health problems therefore requires a thorough understanding and study of population characteristics.
-
-We can also see that in the data table @fig-dataset-overview for 2022, COVID-19 ranks fourth, with 1,547 deaths.
-Alberta's cause-of-death data proves that the coronavirus pandemic is one of the most serious public health crises of the early 21st century, with global implications.
-This figure not only speaks to the outbreak's death rate, but also suggests that the region's health care system may be under strain.
-The COVID-19 pandemic has disrupted the distribution of deaths from previously common causes, which may include some chronic diseases that have long held the top spot, such as cancer and heart disease, even though it was not the leading cause of death in the years shown in the table.
-
-As large datasets are difficult to observe, this report will analyze only specific aspects.
-The original dataset contains eight different causes of death.
-Only the first five causes will be analyzed in this paper.
-
-For simplicity we restrict ourselves to the five most common causes of death in 2022 of those that have been present every year.
 ```{r}
 #| label: tbl-top-five
-#| tbl-cap: Top Five Leading causes of death in 2022
 #| echo: false
@@
 
-head(alberta_cod_top_five)
+alberta_cod_top_five <- data.frame(CauseOfDeath = alberta_cod_top_five)
+
+# New names for the causes of death
+new_names <- c("Organic dementia",
+               "All other forms of chronic ischemic heart disease",
+               "Other ill-defined and unknown causes of mortality",
+               "COVID-19, virus identified",
+               "Malignant neoplasms of trachea, bronchus and lung")
+
+# Assign these new names to the CauseOfDeath column
+alberta_cod_top_five$CauseOfDeath <- new_names
+
+# View the updated data frame
+
+# Create the table with complete variable names
+kable(alberta_cod_top_five, 
+      caption = "Top Five Causes of Death in Alberta for 2022")
+
 ```
+
+We can also see that in the data table for 2022, COVID-19 ranks fourth, with 1,547 deaths.
+Alberta's cause-of-death data proves that the coronavirus pandemic is one of the most serious public health crises of the early 21st century, with global implications.
+This figure not only speaks to the outbreak's death rate, but also suggests that the region's health care system may be under strain.
+The COVID-19 pandemic has disrupted the distribution of deaths from previously common causes, which may include some chronic diseases that have long held the top spot, such as cancer and heart disease, even though it was not the leading cause of death in the years shown in the table.
 
@@
 
-Based on our discussion in the model set-up section, the negative binomial distribution seems more suitable for our research objectives than the Poisson distribution.
-The Poisson model assumes that the mean and variance are equal, but in the real world, they are rarely equal.
-As @tbl-mean-var shown below indicates, the mean and variance are not equal.
-The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.
-
 ```{r}
 #| label: tbl-mean-var
-#| tbl-cap: Comparison of Mean and Variance of Total Death in from 2001 to 2022
 #| echo: false
@@
 # Create a table
-kable(summary_stats, caption = "Comparison of Mean and Variance of Total Deaths")
+kable(summary_stats, caption = "Comparison of Mean and Variance of Total Deaths in Alberta from 2001 to 2022")
 
 ```
+
+Based on our discussion in the model set-up section, the negative binomial distribution seems more suitable for our research objectives than the Poisson distribution.
+The Poisson model assumes that the mean and variance are equal, but in the real world, they are rarely equal.
+As @tbl-mean-var shown below indicates, the mean and variance are not equal.
+The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.
 
@@
 
-We created a figure @fig-ppcheckandposteriorvsprior that includes both the Poisson and negative binomial models.
+We created @fig-ppcheckandposteriorvsprior that includes both the Poisson and negative binomial models.
 For the Poisson model, the actual observed values y are represented in black.
-The predicted values $y_{Rep}$ are shown in light blue, which are replicated samples generated based on the posterior distribution of model parameters.
+The predicted values *$y_{Rep}$* are shown in light blue, which are replicated samples generated based on the posterior distribution of model parameters.
 
@@
 
-For the negative binomial models, the actual observed values y are also represented in black, and the predicted values yrep are likewise shown in light blue, generated based on the posterior distribution of the model parameters.
+For the negative binomial models, the actual observed values `y` are also represented in black, and the predicted values yrep are likewise shown in light blue, generated based on the posterior distribution of the model parameters.
 Compared to the Poisson model, the predictive distributions generated by the negative binomial model appear to be more concentrated, meaning they are more closely clustered around the actual observed values at higher counts, indicating that the negative binomial model may be more suitable for these data.
@@
 #| message: true
-#| warning: true
-#| label: fig-LOOIC
-#| fig-cap: "Comparing the LOO-CV scores"
+#| warning: false
+#| label: tbl-LOOIC
 
+# Perform LOO-CV (Leave-One-Out Cross-Validation) using the loo package
 poisson <- loo(cause_of_death_alberta_poisson, cores = 1)
@@
 
-loo_compare(poisson, neg_binomial)
+# Compare the models
+comparison <- loo_compare(poisson, neg_binomial)
+
+# Convert the comparison to a data frame for better handling (optional)
+comparison_df <- as.data.frame(comparison)
+
+# Select only the columns for ELPD Diff and SE Diff
+selected_comparison_df <- comparison_df[, c("elpd_diff", "se_diff")]
+
+# Set the row names to more descriptive labels
+rownames(selected_comparison_df) <- c("Negative Binomial Model", "Poisson Model")
+
+# Optionally, rename the columns for clarity
+names(selected_comparison_df) <- c("Difference in Expected Log Predictive Density (ELPD Difference)", "Standard Error of ELPD Difference (SE Difference)")
+
+# Now create the table with the updated column names and selected columns
+kable(selected_comparison_df, caption = "LOO-CV Model Comparison: ELPD Difference and SE Difference", align = 'c')
 ```
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-\newpage
-
 # References

### 2024-03-17 paper/paper.qmd
Updated result section

@@
 
-To demonstrate that the negative binomial model is a better fit, we compared the Expected Log Predictive Density (ELPD) difference and the Standard Error (SE) difference between these two models.
+To demonstrate that the negative binomial model is a better fit, we compared the Expected Log Predictive Density (ELPD) difference and the Standard Error (SE) difference between these two models in @tbl-LOOIC.
 The ELPD difference column compares the difference in expected log pointwise predictive density between the two models—the negative binomial model and the Poisson model.
@@
 # Results
+
+## data result
+@fig-planes shows the annual number of deaths from the top five causes of death in Alberta in 2022 and traces the trend from 2000 to 2022.
+In this case, the number of deaths from all other forms of chronic ischemic heart disease shows a relatively stable trend line with a slight increase over time.
+This could mean that the number of deaths from chronic ischemic heart disease remains high, and this slight increasing trend could be related to population growth and aging, as ischemic heart disease is more commonly seen in older adults.
+Beginning in 2020, a trend line in COVID-19 deaths emerges abruptly, coinciding with a pandemic outbreak.
+This trend line, although short-lived, shows a sharp rise in deaths due to the virus, which is consistent with what happened during the pandemic.
+Factors such as vaccination and treatment measures have led to a significant reduction in this trend over time.
+For deaths from malignant tumors of the trachea, bronchus, and lungs, the trend line is relatively flat, suggesting that the number of deaths from these cancers has remained consistent over the years without significant increases or decreases.
+This stable trend may indicate that effective treatments have been able to counterbalance the effects of the increase in cases or that the promotion of regular medical check-ups has helped to reduce mortality from these diseases.
+The trend line for dementia deaths shows a gradual increase in deaths over time, which may be related to the growth of the older population.
+The increase in dementia-related deaths may be due to a number of factors, including increased life expectancy, which has led to a higher proportion of older people in the population, and the lack of effective treatments to prevent the progression of the disease.
+For the "other unspecified and unknown causes" category, there is a sharp increase from 2018 to 2019 and a rapid decline from 2021 to 2022.
+This increase coincided with the period of the COVID-19 pandemic, which we hypothesize may have been due to the fact that the new coronavirus and its variants were not immediately recognized or adequately documented at that time, resulting in many deaths that could not be accurately attributed to known causes.
+In addition, the healthcare system may have been under heavy pressure during the New Crown epidemic and lacked the resources to conduct thorough investigations, which in turn led to a large number of deaths being categorized as unknown causes.
+
 ```{r}
@@
 
-```{r}
-#| label: fig-summary_table
-#| echo: false
-#| tbl-cap: Summary by cause of the number of yearly deaths in Alberta, Canada
-#| warning: false
-#| message: false
-# Calculate summary statistics for total_deaths
-summary_stats <- data.frame(
-  Min = min(alberta_cod$total_deaths, na.rm = TRUE),
-  Mean = mean(alberta_cod$total_deaths, na.rm = TRUE),
-  Max = max(alberta_cod$total_deaths, na.rm = TRUE),
-  SD = sd(alberta_cod$total_deaths, na.rm = TRUE),
-  Var = var(alberta_cod$total_deaths, na.rm = TRUE),
-  N = sum(!is.na(alberta_cod$total_deaths))
-)
-
-# Use kable from knitr to create a table
-kable(summary_stats)
-```
-
-Our results are summarized in @tbl-modelresults.
+## model result
+@tbl-modelresults shows the four most significant causes of death in Alberta from 2001-2022: organic dementia, COVID-19, malignant neoplasms of the trachea, bronchus, and lung, and other ill-defined and unknown causes of mortality.
+The coefficients indicate the relationship between the total number of deaths and the causes of death listed in the table.the model estimates the coefficients of each variable, quantifying the expected change of one unit change in the predictor variable (specific cause of death) against the response variable (in this case, the number of deaths), while holding all other variables constant.
+For predictive modeling, the fit of the model to historical data is crucial because it tells us how well the model explains changes in the data.
+The better the fit, the more confident we can use the model to predict the future.
 
@@
 ```
+### Poisson distribution model
+In the Poisson model, the expected logarithmic function $log(\mu_i)$ is defined as:
+
+$$
+log(\lambda) = 7.484 - 0.152 \cdot COVID - 0.223 \cdot MaligantNeoplasms 
+- 0.400 \cdot OrganicDementia - 0.007 \cdot UnknownCauses
+$$
+
+The intercept represents the expected log count of deaths when all causes are absent.
+In the Poisson model, the baseline number of deaths equals 7.484.
+For COVID-19, the coefficient is -0.152, indicating that the presence of this cause is associated with a slight decrease in the total death count.
+Moreover, for malignant neoplasms with a coefficient of -0.223, suggesting a slightly larger decrease in total death counts when malignant neoplasms are a cause.
+In terms of organic dementia, a coefficient of -0.400 suggests a more substantial decrease in total death counts with organic dementia as a cause.
+For other ill-defined and unknown causes, the coefficient is -0.007, indicating almost no change in the total death count when these are listed as the cause.
+
+### Negative binomial distribution model
+
+In the negative binomial model, the expected logarithmic count is defined as: 
+$$
+log(\mu_i) = 7.482 - 0.129 \cdot COVID - 0.220 \cdot MaligantNeoplasms 
+- 0.396 \cdot OrganicDementia + 0.017 \cdot UnknownCauses
+$$
+
+For a negative binomial model, the baseline number of deaths when all causes are absent equals 7.482.
+For the cause of Covid, each additional count is correlated with a 0.129 decrease in the log count of deaths.
+This also suggests that areas with one more unit of covid-19 reported deaths are expected to see a decrease in the total number of deaths (logarithmic scale).
+In addition, for malignant neoplasms, the coefficient indicated each additional count of deaths due to this cause is associated with a 0.22 decrease in the log count of total deaths.
+Similarly, the coefficient for cause of organic dementia is equal to -0.396, this indicates that for each additional count of deaths due to this cause is associated with a 0.396 decrease in the log count of total deaths.
+Lastly, the coefficient for other ill-defined and unknown causes is small and positive which suggest that an increase in the number of deaths due to these causes are slightly associated with an increase in the log count of total deaths.
+
+Both models’ coefficients represent that as each of these causes of death become a more prevalent reason for death, the total number of deaths decreases in relation to the baseline level set by the intercept, except for "Other ill-defined and unknown causes," which shows no significant effect.
 

### 2024-03-17 paper/paper.qmd
Working on the discussion section

@@
 ---
-title: "My title"
+title: "Analyzi"
 subtitle: "My subtitle if needed"
@@
 
-# Data {#sec-data}
+# Data
 
@@
 
-## Data Method
+## Data Methods
 In the preliminary phase, we worked on organizing and harmonizing the naming of variables in the dataset to enhance its readability and efficiency of analysis.
@@
 
-## Model setup:
+## Model Setup:
 
-In analyzing the association between the total number of deaths and significant causes of death, we used two different regression models to predict, namely the Poisson distribution and the binomial distribution.
+In analyzing the association between the total number of deaths and significant causes of death, we used two different regression models to predict, namely the poisson distribution and the binomial distribution.
 Both the Poisson and binomial distribution models are regression models used to analyze count data, primarily for finding the relationship between independent variables (predictors) and dependent variables (count outcomes).
-In regression models, the independent variables are independent, while the dependent variables depend on the independent variables.
+In regression models, independent variables (also known as predictors or explanatory variables) influence the outcomes, while dependent variables (also known as response variables) are influenced by these independent variables.
 In our model, the independent variables are the different causes of death mentioned in the table, such as COVID-19, malignant neoplasms, organic dementia, etc.
@@
 At the same time, the Poisson distribution is a regression model specifically used for analyzing count data (where the response variable is non-negative).
-In our study, our response variable is the number of deaths, which is definitely a positive number, making the Poisson distribution suitable for our research.
-However, it also has some shortcomings, such as the problem of overdispersion; if the actual data show that the variance is greater than the mean, the Poisson model may not be applicable, and the model's predictive effect will worsen.
+In our study, we use the number of deaths as the response variable, which is inherently a positive integer, thus justifying the initial suitability of the Poisson distribution for our analysis.
+However, the Poisson model comes with limitations, notably the issue of overdispersion—where the observed variance exceeds the mean.
+In such cases, the Poisson model may become inappropriate, leading to a deterioration in the model’s predictive performance.
 
@@
 
-In our data situation, the negative binomial model seems more appropriate than the Poisson model, possibly due to overdispersion in the data.
+In our data situation, the negative binomial model seems more appropriate than the Poisson model, due to overdispersion in the data.
 Therefore, in analyzing causes of death (such as COVID-19, malignant neoplasms, organic dementia, etc.), choosing the negative binomial model may provide more accurate estimates.
 
-Poisson model:
+Our specified model for Poisson is:
 \begin{align} 
@@
 \end{align}
+where $\lambda$ represents the expected rate of occurrence of deaths, $\beta_0$ is the intercept, and $\beta_1$ is the coefficient for the variable $x_1$.
 
-Negative binomial model:
+Our specified model for Negative binomial model is:
 \begin{align} 
@@
 \end{align}
-
-From this formula, we can consider $y_i$ to be the total number of deaths, with the i-th observation representing the cause of death.
-$\beta_0$ represents the intercept, which is the expected log count of the response variable when all predictor variables are held at zero.
+From this formula, we can consider $\mu_i$ to be the expected rate of occurrence of deaths.
+$\beta_0$ represents the intercept, which is the expected log count of the response variable when all predictor variables are held at zero, and $\beta_1$ is the coefficient for the variable $x_1$.
 The intercepts of the two models are similar.
@@
 
-## Model justification
+## Model Justification
 
@@
 
-We created @fig-ppcheckandposteriorvsprior that includes both the Poisson and negative binomial models.
-For the Poisson model, the actual observed values y are represented in black.
+We created @fig-ppcheckandposteriorvsprior that includes both the poisson and negative binomial models.
+For the poisson model, the actual observed values y are represented in black.
 The predicted values *$y_{Rep}$* are shown in light blue, which are replicated samples generated based on the posterior distribution of model parameters.
 
-It can be seen that there is a significant overlap between many of the predictive distributions and the actual observed values, but at higher values, the range of the predictive distribution becomes very wide, indicating that the model may not fit well for high-count values.
+It can be seen that there is a significant overlap between many of the predictive distributions and the actual observed values, but at higher values, the range of the predictive distribution becomes significantly wide, indicating that the model may not fit well for high-count values.
 
-For the negative binomial models, the actual observed values `y` are also represented in black, and the predicted values yrep are likewise shown in light blue, generated based on the posterior distribution of the model parameters.
-Compared to the Poisson model, the predictive distributions generated by the negative binomial model appear to be more concentrated, meaning they are more closely clustered around the actual observed values at higher counts, indicating that the negative binomial model may be more suitable for these data.
-From the figure, it can be observed that the predictive distributions of the negative binomial model match the actual observed values better than those of the Poisson model.
-In the Poisson model's figure, the predicted values align well with the actual values near the peak in the middle but are wider in the tail distribution, which may mean that the Poisson model struggles with extreme values.
-In contrast, in the figure for the negative binomial model, the predicted values seem to follow the actual observed values more closely across the entire range, typically indicating it can better handle the overdispersion of the data.
+For the negative binomial model, similar to the poisson model, the actual observed values `y` are also represented in black, and the predicted values $y_{rep}$ are likewise shown in light blue, generated based on the posterior distribution of the model parameters.
+Compared to the poisson model, the predictive distributions generated by the negative binomial model appear to be more concentrated, meaning that they are more closely clustered around the actual observed values at higher counts, indicating that the negative binomial model may be more suitable for these data.
+From the figure, it can be observed that the predictive distributions of the negative binomial model match the actual observed values better than those of the poisson model.
+In the poisson model's figure, the predicted values align well with the actual values near the peak in the middle but are wider in the tail distribution, which may mean that the Poisson model struggles with extreme values.
+In contrast, in the figure for the negative binomial model, the predicted values seem to follow the actual observed values more closely across the entire range, typically indicating that it can better handle the overdispersion of the data.
 
@@
 
-To demonstrate that the negative binomial model is a better fit, we compared the Expected Log Predictive Density (ELPD) difference and the Standard Error (SE) difference between these two models in @tbl-LOOIC.
-The ELPD difference column compares the difference in expected log pointwise predictive density between the two models—the negative binomial model and the Poisson model.
+To demonstrate that the negative binomial model is a better fit, we compared the Expected Log Predictive Density (ELPD) difference and the Standard Error (SE) difference between these two models.
+As in @tbl-LOOIC, the ELPD difference column compares the difference in expected log predictive density between the two models—the negative binomial model and the Poisson model.
 ELPD is an indicator of model predictive performance, with higher ELPD values typically indicating better model predictive performance.
@@
 
-## data result
+## Data Results
 @fig-planes shows the annual number of deaths from the top five causes of death in Alberta in 2022 and traces the trend from 2000 to 2022.
 In this case, the number of deaths from all other forms of chronic ischemic heart disease shows a relatively stable trend line with a slight increase over time.
-This could mean that the number of deaths from chronic ischemic heart disease remains high, and this slight increasing trend could be related to population growth and aging, as ischemic heart disease is more commonly seen in older adults.
+This could mean that the number of deaths from chronic ischemic heart disease remains high, and this slight increasing trend could be related to population growth and aging, as ischemic heart disease is more commonly seen in elders.
 Beginning in 2020, a trend line in COVID-19 deaths emerges abruptly, coinciding with a pandemic outbreak.
@@
 
-## model result
+## Model Results
 @tbl-modelresults shows the four most significant causes of death in Alberta from 2001-2022: organic dementia, COVID-19, malignant neoplasms of the trachea, bronchus, and lung, and other ill-defined and unknown causes of mortality.
-The coefficients indicate the relationship between the total number of deaths and the causes of death listed in the table.the model estimates the coefficients of each variable, quantifying the expected change of one unit change in the predictor variable (specific cause of death) against the response variable (in this case, the number of deaths), while holding all other variables constant.
+The coefficients indicate the relationship between the expected total number of deaths and the causes of death listed in the table.
+The model estimates the coefficients of each variable, quantifying the expected change of one unit change in the predictor variable (specific cause of death) against the response variable (in this case, the number of deaths), while holding all other variables constant.
 For predictive modeling, the fit of the model to historical data is crucial because it tells us how well the model explains changes in the data.
@@
 ```
-### Poisson distribution model
+### Poisson Distribution Model
 In the Poisson model, the expected logarithmic function $log(\mu_i)$ is defined as:
@@
 
-### Negative binomial distribution model
+### Negative Binomial Distribution Model
 
@@
 
-## Addressing Public Health Challenges in Alberta: Strategies for Health Policy and Social Regulation
+## Environmental impact
+The high incidence of respiratory illnesses in Alberta may be related to the region's unique environmental factors, as Alberta is extremely well developed in the field of source energy, which is also a major source of the province's economy.
+The region is home to a number of similar sectors, and it's possible that working in these industries would expose employees to dangerous compounds, which might lead to an increase in airborne pollutants.
+A large portion of the year is spent in a chilly, dry environment in Alberta.
+Heating systems are often used by citizens, which may cause issues with indoor air quality.
+Heating systems have the potential to worsen respiratory issues by circulating dust, germs, and other small particles through the air and making it drier inside.
+The prevalence of smoking and urbanization-related factors like building development and automobile emissions may potentially have a detrimental effect on respiratory health.
+When taken into account collectively, these variables could put inhabitants of Alberta at greater risk of dying from respiratory diseases.
+All things considered, a complex interplay of variables contributes to Albertans' struggles with respiratory diseases.
+This necessitates a comprehensive public health plan that lowers emissions from vehicles and industry, enhances indoor air quality, and increases public knowledge of healthy lifestyle choices in order to lower the prevalence of respiratory illnesses and the related mortality risk.
+
+## Weaknesses, bias, and ethics
+Future counts may be impacted by abrupt changes in the population or public health crises.
+For instance, the sudden emergence of the coronavirus in a specific year led to numerous fatalities.
+Such unpredictable events can profoundly affect the accuracy of our established models, resulting in erroneous forecasts.
+As a result, models should be updated often with fresh data even if they might shed light on future patterns based on historical data.
+Furthermore, the data's reach is restricted to Alberta, and because different areas and nations have differing degrees of economic strength and health promotion, the data is not indicative of the broader population.
+The influence of regional variations on forecasts should be examined using multilevel modeling or by collecting a larger geographic sample.
+Sociodemographic factors also have a substantial impact on the cause of death statistics.
+The age distribution may be more clearly seen since children and the elderly are more likely than other age groups to suffer from certain diseases.
+It's possible that women are more prone to have certain malignant tumors than males are to suffer from heart disease.
+[@regitz2023gender] Because certain diseases are more frequent in one sex than the other, the sex ratio may thus also have an impact on the cause of mortality.
+Furthermore, racial disparities could influence the likelihood of contracting a disease.
+For instance, genetic variables may be linked to a greater frequency of a certain chronic illness in a given ethnic group.
+Finally, there may be privacy issues involved in collecting these data, and researchers should obtain informed consent from participants.
+
+## Next steps
+Addressing Public Health Challenges in Alberta: Strategies for Health Policy and Social Regulation
 
@@
 Additionally, promoting healthy lifestyles, early cancer screening programs, and providing high-quality cancer treatment services are also crucial [@eastman2023nci].
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
 \newpage

### 2024-03-17 paper/paper.qmd
Updated discussion section

@@
 abstract: "We examined the distribution of the top 30 causes of death for each year between 2001 and 2022 in Alberta, Canada.
-The conifer model and negative binomial regression were used to analyse the long-term leading causes of death and the sudden emergence of specific causes of death.
+The poisson model and negative binomial regression were used to analyse the long-term leading causes of death and the sudden emergence of specific causes of death.
 According to our findings, negative binomial regression improves our ability to predict outcomes when the data is too spread out by fitting the data more accurately.
@@
 
-## Environmental impact
+## Overview of the Finding
+
+For the 2022 data, the process of utilizing different statistical models applied to the historical data to estimate model parameters is used to understand how each cause contributes to the total number of deaths.
+Using the estimated coefficients, the relationship between each cause of death and the total number of deaths can be extrapolated to predict future trends in causes of death.
+
+## Covid Impact
+The data show the severity and mortality of the novel coronavirus during the pandemic.
+The rapid spread of the virus and the lack of treatment and early prevention efforts led to a serious health disaster.
+The lethality of the new coronavirus is reflected not only in the number of deaths directly related to the virus, but also in the increase in unspecified causes of death due to the strain the virus has placed on health care resources.
+Alberta has responded to the epidemic and its impact on mortality by taking action to prevent the spread of the virus.
+In particular, a mass vaccination program has been implemented to reduce the number of severe cases and deaths.
+Preventing the recurrence of such epidemics requires lessons learned from sudden-onset events and stronger response systems.
+This will require greater interregional and international cooperation to improve the responsiveness of public health systems and the ability to rapidly allocate resources in emergencies.
+In addition, the exchange of outbreak data in an open and transparent manner is critical to stopping such outbreaks in the future.
+By allowing Governments and the public to make informed decisions based on the most up-to-date knowledge, such data-sharing not only accelerates scientific understanding of the characteristics of the virus and its mode of transmission, but also reduces mortality rates.
+
+## Environmental Impact
 The high incidence of respiratory illnesses in Alberta may be related to the region's unique environmental factors, as Alberta is extremely well developed in the field of source energy, which is also a major source of the province's economy.

### 2024-03-17 paper/paper.qmd
Final version of paper

@@
 ---
-title: "Analyzi"
-subtitle: "My subtitle if needed"
+title: "Deciphering Mortality Patterns: A Comprehensive Study of Causes of Death in Alberta (2001-2022)"
+subtitle: "Evaluating Predictive Models and Public Health Implications"
 author: 
@@
 
-The mortality rate, often referred to as the death rate, represents an approximation of the fraction of a population that dies within a given time frame [@porta2014dictionary].
-Mortality rates can serve as a crucial indicator of a population’s health status, and it also reveals the impact of diseases and other health-related issues over a period of time.
-This paper explores the leading causes of death in Alberta for crafting effective public health strategies and policies and understanding the most significant health threats affecting a population for researchers[@datasetCOD].
-By identifying the main causes of mortality, health authorities can prioritise research funding towards diseases and conditions that have the highest impact on community health and lifespan [@vargas2019assessment].
+One of the most populated provinces in Canada, Alberta is distinguished for its diverse population and rapid population increase.
+Since population growth has been a major factor in Alberta's economic growth, it is imperative.
+Therefore, for population expansion, it is essential to comprehend the causes of death in this province and to take appropriate action.
+The mortality rate, sometimes called the death rate, is an estimate of the percentage of a population that passes away in a specific amount of time [@porta2014dictionary].
+Over time, mortality rates can provide important insights into the health state of a community by highlighting the effects of various diseases and health-related concerns.
+The government may facilitate public health planning and policy making by identifying health priorities and determining how best to distribute resources through the investigation of death rates.
 
-As discussed in the data section, we used data from Service Alberta [@datasetCOD] on the leading causes of deaths, in which the five most significant causes in 2022 were analysed.
-These five causes are Organic dementia, Other causes not clearly defined, COVID-19, and Cancers of the trachea, bronchus, and lungs.
-It was noted that, among the examples mentioned, the negative binomial regression is more accurate compared to the Poisson model, while Poisson regression is prone to errors.
+In order to create efficient public health plans and policies and to comprehend the most important health hazards impacting a community for academics, this study examines the patterns of the five leading causes of death over a 20-year period.
+As mentioned in the data section, we first provide the eight top causes of death in Alberta in 2022, analyzing the five most important causes as estimands in 2022.
+We do this by using data on fatalities from Service Alberta [@datasetCOD].
+These five reasons include COVID-19, malignant neoplasms of the trachea, bronchus, and lungs, organic dementia, and other causes not fully understood.
+Of them, malignant neoplasms of the trachea have the least effect on the number of fatalities, whereas organic dementia has the most.
+
+We used the Poisson model and the negative binomial model as our two statistical models in the model section.
+In order to provide accurate evaluations of mortality occurrences within a population, the Poisson and negative binomial models are utilized due to their usefulness in assessing count data and allowing overdispersion, respectively.
+We may infer the patterns of these five causes over the last 20 years by determining if the overall number of fatalities and the causes of death are positively or negatively connected.
+Because of the overdispersion in the data, it was shown that the negative binomial regression is more accurate than the Poisson model.
+
+In the result section, we go over the data and model findings.
+We discovered that, with the exception of "Other ill-defined and unknown causes," which shows no significant influence, a rise in the prevalence of each cause of death resulted in a decrease in the overall number of fatalities when compared to the baseline.
+In the discussion section, we address the effects of COVID-19 and environmental variables, as well as outline the study's possible shortcomings and suggest future directions for development.
+
+We hope that by examining the top five causes of death in our article, the public and the Alberta government will recognize the significance of public health and take steps to improve individual health and lower mortality rates.
 
@@
 
-Our data is of leading causes of death (@fig-dataset-overview), from @datasetCOD.
-
 ## Data Sources
 Our research team conducted an in-depth analysis of mortality trends in Alberta, choosing a dataset published on the Alberta Open Government Portal as the basis for their study.
-This dataset was made available to the public beginning in 2015 and was last updated in 2023, and is designed to promote public engagement by providing transparent access to key health statistics.This dataset provides a comprehensive view of deaths across the province, covering the top 30 leading causes of death in each annual count.
+This dataset was made available to the public beginning in 2015 and was last updated in 2023 from @datasetCOD, and is designed to promote public engagement by providing transparent access to key health statistics.
+This dataset provides a comprehensive view of deaths across the province, covering the top 30 leading causes of death in each annual count.
 This dataset was chosen because it provides key statistics on the number of deaths and mortality rates.
@@
 `Accidental poisoning by and...`, 
-`Other chronic obstructive pulmonary diseases`, 
+`Other chronic obstructive pulmonary diseases`. 
 *Ranking*: This is a ranking by the number of deaths caused by each disease, with 1 being the highest.
@@
 
-## Overview of the Finding
-
 For the 2022 data, the process of utilizing different statistical models applied to the historical data to estimate model parameters is used to understand how each cause contributes to the total number of deaths.
